---
title: "Web Scraping Tutorial"
author: "Anri Smith"
date: "16/08/2017"
output: pdf_document
---

# Web Scraping to assist car purchases
The tutorial below describes how to scrape information from the \textit{www.cars.co.za} website. In this particular example information with regards to the sale of the Audi A3 model in the Western Cape is of interest.

## STEP 1
Firstly the following packages need to be installed/loaded.
\bigskip
```{r setup, include=TRUE, echo=TRUE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(rvest)
library(tidyverse)
library(stringr)
```

## STEP 2

In the web-browser one can now navigate to the \textit{www.cars.co.za} page and select the province and car model you are interested in (The Western Cape and the Audi A3 in this example). Once this page is reached, copy the URL so that a connection can be formulated to this page using the \textit{read\_html} function.
\bigskip
```{r main page, echo = T, results = "show" }
#-----the first summary page
Audia3.cars <- read_html("http://www.cars.co.za/usedcars/Western-Cape/Audi/A3/")
Audia3.cars
```
\bigskip
One will note that there are links to more pages displaying cars that fit the specifications given. Access to these pages is also required. All of the links on the current page can be viewed, labeled \textit{page\_links} in this case. From here the links to the other summary pages can be extracted the same way as before. 
\bigskip
```{r other summaries , echo = T, results = "show" }
#-----all links on first summary page
page_links <- Audia3.cars %>% html_nodes("a") %>% html_attr("href") 
#print(page_links)

#-----extracting links to other summary pages
car_pages <- str_subset(page_links,
                        "(http://www.cars.co.za)*(usedcars)*(Western-Cape)*(Audi)*(A3).*(P=).")
car_pages
car_pages.new <- c()
for (i in 1:length(car_pages))
{
  tmp <- cbind("http://www.cars.co.za", car_pages[i])
  car_pages.new[i] <- str_c(tmp, collapse = "")
}
```

## STEP 3

Now that links have been established to all of the relevant pages, one wishes to extract more information on each car that is advertised. In order to do this, links to these indiviual advertisements need to be made. Once this has been done, CSS selectors can be used to extract the relevant information on each car. 
\bigskip

```{r full ad , echo = T, results = "show" }

#-----extracting the links to full ads on each summary page.
car_links <- c()
for(i in car_pages.new){
  summary_i <- read_html(i)
  car_i_links <- summary_i %>% html_nodes("a") %>% html_attr("href") 
  car_links_i <- str_subset(car_i_links,"(http://www.cars.co.za)*(for-sale).")
  car_links <- c(car_links, car_links_i)
}
car_links <- sample(unique(car_links))
```

## STEP 4
One will now need to navigate through the browser to any car advert on the website that meets the search specifications. Here one will make use of the CSS selector (This tool can be downloaded from \textit{selectorgadget.com\/}). The tool essentially allows you to highlight the information of interest on the web page. In order to extract a given piece of information the tag associated with the CSS selector for that information must be copied in as shown below in the \textit{html\_nodes} function.

Once all the information of interest has been collected the data can be placed in a data frame as per usual.

NOTE: This step has only been performed on three advertisements for computational reasons. If one has the patience the number of entries can be increased.
\bigskip
```{r css selectors , echo = T, results = "show" }
car_data <- data.frame()
for(i in car_links[1:3])
{
  #car html
  car <- read_html(paste("http://www.cars.co.za",i, sep=""))
  
  #car details
  model <- car %>% html_nodes(css = ".heading_size_xl") %>% 
    html_text(trim = TRUE)
  price <- car %>% html_nodes(css = ".vehicle-view__price") %>%
    html_text(trim = TRUE)
  details.label <- car %>% html_nodes(css = ".vehicle-details__label") %>% 
    html_text(trim = TRUE)
  details <- car %>% html_nodes(css = ".vehicle-details__value") %>%
    html_text(trim = TRUE)
  
  
  model <- ifelse(length(model) > 0, model, NA)
  model <- strsplit(model, "Audi A3")[[1]][2]
  model <- strsplit(model, "for sale")[[1]][1]
  
  price <- ifelse(length(price) > 0, price, NA)
  cbind(details.label, details)
  
  #preparing current entry
  this_car <- data.frame(model = model, price = price, mileage = details[1], 
                         transmission = details[2],   year = details[3], 
                         fuel.type = details[4], Condition = details[5], 
                         Area = details[6], Colour = details[7], 
                         Reference = details[8])
  
  car_data <- rbind.data.frame(car_data,this_car)
  
  Sys.sleep(sample(seq(10, 30, by=1), 1))
  
  }
```

## STEP 5
The data can now easily be viewed in a table to assist one in making an informed purchase.
\bigskip
```{r data , echo = T, results = "show" }
as.tibble(car_data)


```










