---
title: "R Notebook"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      tidy.opts=list(width.cutoff=60),
                      tidy=TRUE)
```

```{r}

rm(list = ls())

suppressMessages(library(rvest))
suppressMessages(library(tidyverse))
suppressMessages(library(stringr))
```

# Introduction

The goal of this script is to retrieve information about cars from Autortrader.co.za in order to analyse it.

# Scraping

Because there are over 60 000 vehicles listed on Autotrader.co.za, only one vehicle brand will be investigated - in this case "Hyundai":

```{r}
brand_select <- "hyundai"
```

After some investigation, the GET request to retrieve the search page with the most results for a specific brand was discovered. The brand name is substituted into the GET request url:

```{r}
search_page_0_address <- paste0("https://www.autotrader.co.za/makemodel/make/", toupper(brand_select), "/perpage/60/search")
```

From the search page, retrieve all the links to individual vehicles:

```{r}

#Read the search page
search_page_0 <- read_html(search_page_0_address)

# select all "a" tags with class "mobile-cav-link" (add #id_name before tag.name as an option)
search_page_0_elements <- html_nodes(x = search_page_0, "a.mobile-cav-link")
car_links <- search_page_0_elements %>% html_attr("href")

last_page <- search_page_0 %>% html_node("li.to-last-page") %>% html_attr("data-page") %>% as.numeric()

```

Once we have links to all the vehicles on the first search page, set up a loop in order to collect information from each individual vehicle's page:

```{r}

#####
page_data <- read_html(car_links[1])

vehicle_data <- rbind(
    data.frame("spec" = page_data %>% html_nodes("div.spec-name") %>% html_text(),
             "val" = page_data %>% html_nodes("div.spec-value") %>% html_text()),
    data.frame("spec" = page_data %>% html_nodes("span.label") %>% html_text(),
             "val" = page_data %>% html_nodes("span.value") %>% html_text()),
    data.frame("spec" = "Price",
             "val" = page_data %>% html_nodes("div.price") %>% html_nodes("h2") %>% html_text()),
    data.frame("spec" = "URL",
             "val" = car_links[1])
  )

all_veh_data <- data.frame(t(vehicle_data))
colnames(all_veh_data) <- unlist(all_veh_data[1,])
all_veh_data <- all_veh_data[-1,]

### now get the info for the rest of the links

# for (i in 2:length(car_links)){
for (i in 2:60){

  link <- car_links[i]
  page_data <- read_html(link)
  
  vehicle_data <- rbind(
    data.frame("spec" = page_data %>% html_nodes("div.spec-name") %>% html_text(),
             "val" = page_data %>% html_nodes("div.spec-value") %>% html_text()),
    data.frame("spec" = page_data %>% html_nodes("span.label") %>% html_text(),
             "val" = page_data %>% html_nodes("span.value") %>% html_text()),
    data.frame("spec" = "Price",
             "val" = page_data %>% html_nodes("div.price") %>% html_nodes("h2") %>% html_text()),
    data.frame("spec" = "URL",
             "val" = link)
  )
  
  vehicle_data <- data.frame((t(vehicle_data)))
  colnames(vehicle_data) <- unlist(vehicle_data[1,])
  vehicle_data <- vehicle_data[-1,]
  # rownames(vehicle_data)[1] <- link
  
  all_veh_data <- suppressWarnings(suppressMessages(bind_rows(all_veh_data, vehicle_data))) #is this better than bind_rows
  
  #TODO: add timing to ensure closing of dead connections
  
}
```

Go further by collecting all possible links to the vehicle brand, and then iterating through that in the same way:

```{r}

# for (i in 2:last_page){
for (i in 2:4){
  search_page_address <- paste0("https://www.autotrader.co.za/makemodel/make/", toupper(brand_select), "/page/", i, "/perpage/60/search")
  #Read the search page
  search_page <- read_html(search_page_address)
  
  # select all "a" tags with class "mobile-cav-link" (add #id_name before tag.name as an option)
  search_page_elements <- html_nodes(x = search_page_0, "a.mobile-cav-link")
  car_links <- c(car_links, search_page_elements %>% html_attr("href"))
}
```

```{r}

# for (i in 61:length(car_links)){
for (i in 61:150){

  link <- car_links[70]
  page_data <- read_html(link)
  
  vehicle_data <- rbind(
    data.frame("spec" = page_data %>% html_nodes("div.spec-name") %>% html_text(),
             "val" = page_data %>% html_nodes("div.spec-value") %>% html_text()),
    data.frame("spec" = page_data %>% html_nodes("span.label") %>% html_text(),
             "val" = page_data %>% html_nodes("span.value") %>% html_text()),
    data.frame("spec" = "Price",
             "val" = page_data %>% html_nodes("div.price") %>% html_nodes("h2") %>% html_text()),
    data.frame("spec" = "URL",
             "val" = link)
  )
  
  vehicle_data <- data.frame((t(vehicle_data)))
  colnames(vehicle_data) <- unlist(vehicle_data[1,])
  vehicle_data <- vehicle_data[-1,]
  
  all_veh_data <- suppressWarnings(suppressMessages(bind_rows(all_veh_data, vehicle_data))) #is this better than full_join?
  
}

```

# Cleaning

Clean important numerical data:

```{r}

all_veh_data_copy <- cbind(all_veh_data)

all_veh_data_copy <- select(all_veh_data_copy,
                       'Registration year',
                       Make,
                       Model,
                       Variant,
                       Mileage,
                       Price,
                       everything())


all_veh_data_copy$Price <- str_replace(all_veh_data_copy$Price, "R", "")
all_veh_data_copy$Price <- str_replace(all_veh_data_copy$Price, " ", "")
all_veh_data_copy$Price <- as.numeric(all_veh_data_copy$Price)

all_veh_data_copy$Mileage <- str_replace(all_veh_data_copy$Mileage, "Km", "")
all_veh_data_copy$Mileage <- str_replace(all_veh_data_copy$Mileage, ",", "")
all_veh_data_copy$Mileage <- str_replace(all_veh_data_copy$Mileage, " ", "")
all_veh_data_copy$Mileage <- as.numeric(all_veh_data_copy$Mileage)

```

# Analysis

Present some outcomes of the data:

```{r}

# Mileage vs. year

plot(all_veh_data_copy$Mileage, all_veh_data_copy$Price,
     main = "Price vs. Mileage",
     xlab = "Mileage",
     ylab = "Price")


# For each year: mileage vs price


# Group by variant - show boxplot



```
