---
title: "Webscraping"
author: "Curtly Blows - BLWCUR001"
date: "August 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
THe aim is to extract the necessary information from this website to use as a reference for a paper or research purposes.

We start by importing the necessary libraries.

```{r, echo=F}
library(rvest)
library(tidyverse)
library(stringr)
```
We then load the web page into a variable
```{r}
web=read_html("https://www.universetoday.com/30713/galaxy-evolution/")
```
We locate and store the different components of the web page. The images can be used to support the information in the text. The actual text which can then be summarised.
```{r}
web_imgs = html_nodes(x = web, css = "img")
web_text=html_nodes(x = web, css = "h2,p")
```
We then store the text in plain format, without html characters and store them in a dataframe for easier interperation
```{r}
text=html_text(web_text, trim = TRUE)
text=as.tibble(text)
```
We can then extract the first paragraph of text and a picture url 
```{r}
text[1]
web_imgs[6]
```

