---
title: "Lesson 3 - Web Scraping"
author: "Delyno du Toit"
date: "8/20/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- Load the libraries-->
```{r}
library(rvest)
library(tidyverse)
library(stringr)
```

Scraping gumtree.co.za for harley-davidson ads
 
Go to gumtree.co.za and click on "Motorcycles & Scooters". On the left sidebar, go to "Make" and select "Harley-Davidson".  This brings up the first page of harley-davidsons for sale.  INcluded in the page at the bottom are the number of pages are links to further pages displaying harley-davidsons for sale.  First we read in the first page and identify all hyperlinks on that page.

```{r read, echo = FALSE, include = TRUE}
hd <- read_html("https://www.gumtree.co.za/s-motorcycles-scooters/harley+davidson/v1c9027a1map1")
hd_links <- hd %>% html_nodes("a") %>% html_attr("href") 
print(hd_links)
```

Next, we need to identify just those hyperlinks that load pages with harley-davidons summaries. Within the url the string "/s-motorcycles-scooters/harley" refers to the page links

```{r pages, , echo = FALSE, include = TRUE}
hd_pages <- str_subset(hd_links,"(/s-motorcycles-scooters/harley).*(page)")
hd_pages <- paste0("https://www.gumtree.co.za", hd_pages)
hd_pages
```

For each of the summary pages, we extract the hyperlinks that lead to the full harley-davison ads.  The url changes now from "/s-motorcycles-scooters/harley" to "/a-motorcycles-scooters/harley"
```{r detail, echo = FALSE, include = TRUE}
hd_links <- c()
for(i in hd_pages){
  hd_i <- read_html(i)
  hd_i_links <- hd_i %>% html_nodes("a") %>% html_attr("href") 
  hd_links_i <- str_subset(hd_i_links,".*a-motorcycles-scooters")
  hd_links <- c(hd_links, hd_links_i)
}
# remove any duplicates and reorder
hd_links <- sample(unique(hd_links))
hd_links
```


Now read each of those pages and extract the data we want. 
```{r dataframe, echo = FALSE, include = TRUE}}
hd_data <- data.frame()
for(i in hd_links){   
  
      # read hd ad html
  hd <- read_html(paste0("https://www.gumtree.co.za",i))
  
      # get the ad text 
  title <- hd %>% html_nodes(css = ".myAdTitle") %>% html_text(trim = T)

    # get hd data
  amount <- hd %>% html_nodes(css = ".amount") %>% html_text(trim = TRUE)
  description <- hd %>% html_nodes(css = ".description") 


    # if couldn't find data on webpage, replace with NA
  amount <- ifelse(length(amount) > 0, amount, NA)
  description <- ifelse(length(description) > 0, html_text(description, trim = TRUE), NA)

  
    # store results
  this_hd <- data.frame(title = title, amount = amount, description = description)
  hd_data <- rbind.data.frame(hd_data,this_hd)
  
hd_data
  
}
```

