---
title: "ScrapingTutorial"
author: "Karsten"
date: "21 August 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      tidy.opts=list(width.cutoff=60),
                      tidy=TRUE)
```

#Web-scraping tutorial

First install all the packages below, do this by RStudio, the R-shell with
the command "install.packages("<wanted package")" or other perfered method.

```{r}
library(rvest)
library(tidyverse)
library(stringr)
```

Loading in the page, in this case the page of Fox News. Doing this by using "read_html",
from the "rvest" package. This takes in and read a webpage.

```{r}
fox_page <- read_html("http://www.foxnews.com/")
```

Now we want to get some of the headlines that are "Breaking News" this day. First we
will look at what is the main headline. The webpage is built up as a CSS (Cascading Style sheet).
This style the page by using codes that style the page in the manner desired.
For example the size of the headings are coded with "h<number>" where <number> is replaced by an
integer with bigger integers meaning smaller headings.
Now we will read the main healine of Fox News at this time. The command "html_nodes" will do this for us.
This takes in a webpage and a css-string. To get what is in the css-string you can use a addon like Selector-
Gagdet in chrome or just inspect the page.

```{r}
fox_main_heading <- html_nodes(x = fox_page, css = "h1 a")
fox_main_heading
```

Note that this may give two or more URLs, this is because more than the main headline use this CSS-coding.
Going back to the CSS-selector noticing that putting on #big-top in the start will fix this.

```{r}
fox_main_heading <- html_nodes(x = fox_page, css = "#big-top h1 a")
fox_main_heading
```

Now we want to see what is in the headline. Doing this by using "html_text". Anabling the trim option to get rid
of unnecessary white spaces.

```{r}
fox_main_heading_text <- html_text(fox_main_heading, trim = TRUE)
fox_main_heading_text
```

Now we may want to have a list of other headlines on Fox News. Using the CSS-selector it seems most other headings
use the h3 CSS-code.

```{r}
fox_headings <- html_nodes(x = fox_page, css = "h3")
fox_headings_text <- html_text(fox_headings, trim = TRUE)
fox_headings_text
```

Under is a function which takes in a list of news headlines and a word and returns all the headlines which contains this word. To do this it iterates through the list and use the function "grepl" to check. The fixed=TRUE is used to avoid signs like + and * to be interpreted as regex expressions.

```{r}
containsWord <- function(
  headings_list,
  word){
    wanted_headings <- c()
    for(i in headings_list){
      if(grepl(word,i,fixed=TRUE)){
        wanted_headings <- c(wanted_headings, i)
      }
    }
    return(wanted_headings)
}
```


Under the function is called to see headings from Fox News containing "Trump".
```{r}
trump_headlines <- containsWord(fox_headings_text, "Trump")
trump_headlines
```

Now we'll make a table over all the matches that are left in the Premier League. First we read the html page from theguardian.
```{r}
matches_page <- read_html("https://www.theguardian.com/football/premierleague/fixtures")
```

Then extract the home and away teams. Doing this by using css coding.
```{r}
matches_homeTeam <- html_nodes(x = matches_page, css = ".football-match__team--home")
matches_awayTeam <- html_nodes(x = matches_page, css = ".football-match__team--away")
```


Making variables on text-form and triming them to get rid of unnecessary lineshits and blank spaces.
```{r}
homeTeams <- html_text(matches_homeTeam, trim=TRUE)
awayTeams <- html_text(matches_awayTeam, trim=TRUE)

homeTeams
awayTeams
```

Then we use the data_frame command to make a table with which teams will be playing.
```{r}
premier_table <- data.frame(homeTeams, awayTeams)
premier_table
```

It might be useful to see the number of times a variable is listed. Then we can make a dataframe like this. Under we list the amount of home matches left for the teams.

```{r}
num_homeMatches <- as.data.frame(table(homeTeams))
num_homeMatches
```

To plot this as a barchart we can use the "barplot" function and specify which axis we want to have the different values on.

```{r}
barplot(num_homeMatches$Freq, names.arg = num_homeMatches$homeTeams, main = "Number of home matches left", xlab="Team", ylab="Matches")
``` 

#Sorry for the names of the teames being to long to represent the bars in a more visually correct. But hope this guide was of some use. 